---
title: 'Assignment 2: HA 6'
author: "Group 1: Andrew Carson, Nathan Cooper, Baron Curtin, Heather Geiger, Alejandro Osborne"
date: "February 2, 2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(suppressWarnings(library(tidyverse)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(fpp2)))
```

## HA 6.2
The plastics data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.
Plot the time series of sales of product A. 

a. Can you identify seasonal fluctuations and/or a trend-cycle?

```{r, include=TRUE}
autoplot(plastics) + xlab("Month") + ylab("1000 $") +
  ggtitle("Monthly Plastic Sales")
```

The data has approximately 1 month (peak to peak) seasonality and an apprearent upward trend, that may be cyclic behavior since this only covers 6 months of data.


b. Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

```{r}
plastics
autoplot(plastics)
plas_trend <- ma(plastics, order = 12, centre = TRUE)
plas_trend
autoplot(plas_trend)
plas_detrend <- plastics/plas_trend
plas_detrend
plas_seas <- colMeans(t(matrix(plas_detrend,nrow = 12)), na.rm = TRUE)
plas_seas
autoplot(as.ts(plas_seas))
random_plas = plastics/ (plas_trend* plas_seas)
(random_plas)
random_plas
```

```{r}
plastics %>% decompose(type="multiplicative") %>%
  autoplot() + xlab("Month") +
  ggtitle("Classical multiplicative decomposition
    of Plastic Sales")
```

```{r}
plastics %>% decompose(type="multiplicative")
```

c. Do the results support the graphical interpretation from part a?

Yes the trend is sloped upward and the seasonal trend reaches maximum once a month.

d. Compute and plot the seasonally adjusted data.

```{r}
plas_seas <- decompose(plastics, type = 'multiplicative')$seasonal
plas_seas

plas_seas_adj = plastics/plas_seas
autoplot(plas_seas_adj)
```

e. Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

```{r}
plastics[30]
new_val <- plastics[30]+500
plas_outlier <- replace(plastics,30, new_val)
plas_outlier[30]

plas_seas_outlier <- decompose(plas_outlier, type = 'multiplicative')$seasonal
plas_seas_outlier
plas_seas_adj_outlier = plas_outlier/plas_seas_outlier
autoplot(plas_seas_adj_outlier)
```

It adds a spike to the seasonally adjusted data where the outlier is.

f. Does it make any difference if the outlier is near the end rather than in the middle of the time series?

```{r}
plastics[5]
new_val <- plastics[5]+500
plas_outlier <- replace(plastics,5, new_val)
plas_outlier[5]

plas_seas_outlier <- decompose(plas_outlier, type = 'multiplicative')$seasonal
plas_seas_outlier
plas_seas_adj_outlier = plas_outlier/plas_seas_outlier
autoplot(plas_seas_adj_outlier)
```

```{r}
plastics[55]
new_val <- plastics[55]+500
plas_outlier <- replace(plastics,55, new_val)
plas_outlier[55]

plas_seas_outlier <- decompose(plas_outlier, type = 'multiplicative')$seasonal
plas_seas_outlier
plas_seas_adj_outlier = plas_outlier/plas_seas_outlier
autoplot(plas_seas_adj_outlier)
```

Yes, the spike occurs where the outlier is.

## HA: 6.6

We will use the bricksq data (Australian quarterly clay brick production. 1956-1994) for this exercise.

a. Use an STL decomposition to calculate the trend-cycle and seasonal indices. (Experiment with having fixed or changing seasonality.)

```{r, include=TRUE}
bricksq %>% autoplot()

bricksq %>%
  stl(t.window=13, s.window="periodic", robust=TRUE) %>%
  autoplot()

bricksq %>%
  stl(t.window=13, s.window=7, robust=TRUE) %>%
  autoplot()

bricksq %>%
  stl(t.window=13, s.window=25, robust=TRUE) %>%
  autoplot()

bricksq %>%
  stl(t.window=13, s.window=81, robust=TRUE) %>%
  autoplot()
```

We see that a high s.window yeilds the same as periodic.

```{r, include=TRUE}
bricksq %>% autoplot()

bricksq %>%
  stl(t.window=13, s.window="periodic", robust=TRUE) %>%
  autoplot()

bricksq %>%
  stl(t.window=25, s.window="periodic", robust=TRUE) %>%
  autoplot()

bricksq %>%
  stl(t.window=7, s.window="periodic", robust=TRUE) %>%
  autoplot()

bricksq %>%
  stl(t.window=3, s.window="periodic", robust=TRUE) %>%
  autoplot()
```

The t.window parameter adjusts the goodness of fit of the trend, low values tend to under fit the trend, and high values tend to over fit the trend.

b. Compute and plot the seasonally adjusted data.

```{r, include=TRUE}
brick_seas <- decompose(bricksq, type = 'multiplicative')$seasonal
brick_seas

brick_seas_adj = bricksq/brick_seas
autoplot(brick_seas_adj)
```

c. Use a naïve method to produce forecasts of the seasonally adjusted data.

```{r, include=TRUE}
brick_naive <- naive(brick_seas_adj, h=12)
brick_naive

autoplot(brick_seas_adj) +
  autolayer(brick_naive,
    series="Naïve", PI=TRUE)
```

d. Use stlf() to reseasonalise the results, giving forecasts for the original data.

```{r, include=TRUE}
fcast <- stlf(brick_seas_adj, method='naive')
autoplot(fcast)
```


e. Do the residuals look uncorrelated?

```{r, include=TRUE}
res_brk <- residuals(brick_naive)
autoplot(res_brk)
gghistogram(res_brk) + ggtitle("Histogram of residuals")
ggAcf(res_brk) + ggtitle("ACF of residuals")

res_fcast <- residuals(fcast)
autoplot(res_fcast)
gghistogram(res_brk) + ggtitle("Histogram of residuals")
```

They seem to increse as time increases, so no they do not look uncorrelated. Also a couple of the lags show significance, that may or may not be by chance.

f. Repeat with a robust STL decomposition. Does it make much difference?

```{r, include=TRUE}
brick_stl <- brick_seas_adj %>% stlf(t.window=7, s.window="periodic", method = 'naive', robust=TRUE) 
brick_stl %>% autoplot()
```

In this instance, it the residuals appear to behave similarly to a STL decompostion that is not robust.

```{r}
res_brk2 <- brick_stl %>% residuals()
res_brk2 %>% autoplot()
gghistogram(res_brk2) + ggtitle("Histogram of residuals")
ggAcf(res_brk2) + ggtitle("ACF of residuals")
```

g. Compare forecasts from stlf() with those from snaive(), using a test set comprising the last 2 years of data. Which is better?

```{r}
train <- bricksq %>% window(1956, c(1992, 3))
test <- bricksq %>% window(c(1992, 4), c(1994, 4))
train_stlf <- train %>% stlf(method='naive', h=8)
train_snaiv <- train %>% snaive(h=8)
autoplot(bricksq) +
  autolayer(train_stlf, series="STL", PI=FALSE) +
  autolayer(train_snaiv, series="Seasonal naïve", PI=FALSE) +
  xlab("Quarter") + ylab("Bricks") +
  ggtitle("Forecasts for quarterly brick production") +
  guides(colour=guide_legend(title="Forecast"))
```

From the graphs, it is hard to tell.

```{r}
accuracy(train_snaiv, test)
accuracy(train_stlf, test)
```

From the accuracy fucnction, it appears that the STL decomposition performs better by yeilding less error in ME, RMSE, MAE, MPE, MAPE, and MASE.